Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
abstract edits done

\begin{abstract}

	Causal discovery has received significant attention in the Directed
	Acyclic Graphs (DAGs) literature, leading to the development of
	numerous automated algorithms for learning DAGs from data. However,
	their adoption in applied domains remain limited, as researchers often
	prefer to construct DAGs manually based on domain knowledge. This
	preference arises due to several practical challenges with automated
	algorithms, such as their tendency to make obvious errors and output
	Markov Equivalence Classes (MECs) instead of a DAG. To assist
	researchers in constructing DAGs manually, we propose an iterative
	structure learning approach that combines domain knowledge with
	data-driven insights. Our method leverages Conditional Independence
	(CI) testing to iteratively identify variable pairs where an edge is
	either missing or superfluous. Based on this information, we can choose
	to add missing edges with appropriate orientation based on domain
	knowledge or remove unnecessary ones. We also give a method to rank
	these missing edges based on their impact on the overall model fit.
	Empirical results show that this iterative approach achieves
	performance comparable to automated algorithms if the orientation of
	new edges is correctly determined in at least one out of two cases.
	Additionally, we show that, in the absence of an expert, a Large
	Language Model (LLM) can be used to determine edge orientations. We
	provide an implementation of our approach as both a web tool: <redacted
	for review> and a Python package <redacted for review>.

\end{abstract}

>>>>

\begin{abstract}
	Numerous causal discovery algorithms were developed to automatically learn
	directed acyclic graphs (DAGs) and other causal models from data. However,
	their adoption in applied domains remains limited, as researchers often
	prefer to construct DAGs manually based on domain knowledge. This
	preference arises due to several practical challenges with automated
	algorithms, such as their tendency to produce results that contradict
	obvious domain knowledge and their inability to distinguish Markov equivalent
	models. To assist researchers in constructing DAGs manually, we propose an iterative
	structure learning approach that combines domain knowledge with
	data-driven insights. Our method leverages conditional independence
	testing to iteratively identify variable pairs where an edge is
	either missing or superfluous. Based on this information, we can choose
	to add missing edges with appropriate orientation based on domain
	knowledge or remove unnecessary ones. We also give a method to rank
	these missing edges based on their impact on the overall model fit.
	In a simulation study, we find that this iterative approach to leverage domain 
	knowledge already starts outperforming purely data-driven structure learning if 
	the orientation of new edge is correctly determined in at least one out of two cases.
	We present a proof-of-concept implementation using a large language 
	model as a domain expert and a graphical user interface designed to 
	assist human experts with DAG construction.
\end{abstract}
